Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-02-01 Multiresolution_framework/ltx/manuscript.tex
minor edit
minor edited to author list
minor edit to author list

\documentclass[review,authoryear,3p]{elsarticle}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{setspace} 
\usepackage[usenames,dvipsnames]{color}

\newcommand{\dean}[1]{\textcolor{green}{#1}}
\newcommand{\parham}[1]{\textcolor{blue}{#1}}
% \newcommand{\mike}[1]{\textcolor{Plum}{#1}}

%\newfloat{algorithm}{htbp}{loa}%[chapter]
%\floatname{algorithm}{Algorithm}
\journal{Neuroimage}
\begin{document}
\begin{frontmatter}
\title{Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model}

\author[US]{P. Aram\corref{cor1}}
\ead{parham.aram@univ-amu.fr}
\author[DEEE,BEI,UE]{D. R. Freestone}
\ead{dfreestone@bionicear.org} 
\author[UE,CU]{M. Dewar}
\ead{mike.dewar@columbia.edu}
\author[UM]{K. Scerri}
\ead{kenneth.scerri@um.edu.mt}
\author[DEEE,BEI]{D. B. Grayden}
\ead{grayden@unimelb.edu.au}
\author[US]{V. Kadirkamanathan}
\ead{visakan@sheffield.ac.uk}

\address[US]{Department of Automatic Control and Systems Engineering, University of Sheffield, Sheffield, UK}
\address[DEEE]{Department of Electrical and Electronic Engineering, University of Melbourne, Melbourne, VIC, Australia}
\address[BEI]{The Bionic Ear Institute, East Melbourne, VIC, Australia}
\address[UE]{Institute for Adaptive and Neural Computation, University of Edinburgh, Edinburgh, UK}
\address[CU]{Department of Applied Physics and Applied Mathematics, Columbia University, New York, NY, USA}
\address[UM]{Department of Systems and Control Engineering, University of Malta, Msida, MSD, Malta}

\cortext[cor1]{Corresponding author}

\begin{abstract}
%\boldmath
We develop a multi-resolution approximation (MRA) framework for the integro-difference equation (IDE) neural field model based on semi-orthogonal cardinal B-spline wavelets. State and parameter estimation is performed using the Expectation Maximization (EM) algorithm. A synthetic example is provided
to demonstrate the framework.
\end{abstract} 
\begin{keyword}
  Neural Field Model, multi-resolution approximation (MRA), Expectation Maximisation (EM) algorithm
\end{keyword}

\end{frontmatter}

\section{Introduction}
The human cerebral cortex has a multi-resolution architecture, where spatial scales for information processing range from ion channels, to single neurons, to networks of millions of neurons. This multi-resolution cortical architecture poses major modeling challenges to efficiently describe the brain's dynamics. This paper introduces a multi-resolution data-driven framework for neural field modeling, the Multi-Resolution Approximate Integro-Difference Equation (MRAIDE), to address this challenge.

@@

\section{IDE Neural Field Model}
\singlespacing
\begin{table*}[!t]
\begin{tabular}{|l|l|l|}
	\hline
	\textbf{Symbol} & \textbf{Quantity} & \textbf{Units}\\
	\hline
	\multicolumn{3}{|c|}{\emph{Domain and indices}}\\
	\hline
	$\Omega$ & spatial domain & n.a.\\
	$\mathbf{r}$ & spatial location & [mm, mm]\\
	$t$ & time & s\\
	\hline
	\multicolumn{3}{|c|}{\emph{Model}}\\
	\hline
    $y_t$ & observation & mV\\
    $v(\mathbf{r},t)$ & mean membrane potential & mV \\
	$f(v\left(\mathbf{r},t\right))$ & firing rate function & spikes.s$^{-1}$\\
	$e(\mathbf{r},t)$ & field disturbance, with covariance function $\eta(\mathbf r-\mathbf r')$ & mV\\
	$\boldsymbol\varepsilon_t$ & observation noise, with covariance matrix $\Sigma_\varepsilon$ & mV\\
	$m(\mathbf{r}_n)$ & observation kernel where $n$ is sensor index $n=1,..,n_y$ & n.a. \\
	$w(\mathbf{r})$ & spatial connectivity kernel & n.a.\\
	$h(t)$ & post-synaptic response kernel & mV\\
	$\tau$, $\xi$ & membrane time constant \& time constant parameter & s, n.a.\\
	$\delta(t)$ & Dirac-delta function & n.a.\\
	\hline
	\multicolumn{3}{|c|}{\emph{Reduced Model}} \\
	\hline
   	$\mathbf{x}_t$ & state vector at time $t$ & mV\\
   	$\mathbf{e}_t$ & state disturbance vector, with covariance $\Sigma_e$ & mV\\
   	$\mathbf{C}$ & observation matrix & n.a. \\
	\hline
	\multicolumn{3}{|c|}{\emph{Frequency Analysis}} \\
	\hline
	$\boldsymbol{\nu}$, $\boldsymbol{\nu}_c$ & spatial frequency and spatial cutoff frequency & cycles/mm \\
	$\sigma_{\nu}^2$ & variance of Fourier transformed Gaussian basis function & (cycles/mm)$^2$\\
	\hline
	\multicolumn{3}{|c|}{\emph{Estimation}} \\
	\hline
	$\hat{\mathbf{x}}_t^{f-}$, $\hat{\mathbf{x}}_t^f$ & forward prior and posterior state estimates & mV\\
	$\hat{\mathbf{x}}_t^{b}$ & backward posterior state estimates & mV\\
	$P^{f-}_t$, $P^f_t$  & forward prior and posterior covariance matrices & mV$^2$\\
	$P^b_t$ & backward posterior covariance matrices & mV$^2$\\
	$M_t$& cross-covariance matrix & mV$^2$\\
	$\mathcal K_{t} $ & filter gain & n.a.\\ 
	$\mathcal S_{t} $ & smoother gain & n.a.\\
	\hline
\end{tabular}
\caption{\textbf{Notation.} The first column of the table gives the symbols for all the notation used in the paper. The second column provides a brief description of the quantity, and the third column provides the SI units where relevant.}
\label{tab:Notation}
\end{table*}  
\doublespacing

@@

The multi-resolution approximation \citep{Mallat1989a} of the neural IDE (MRAIDE) is obtained by decomposing both the field, $v_t(\cdot)$, and the connectivity kernel, $w(\cdot)$, (assuming square-integrable functions) using translations and dilations of a scaling function $\phi(\mathbf{r})$ and a mother wavelet $\psi(\mathbf{r})$. Considering a  one-dimensional field, the static connectivity kernel is decomposed as,
\begin{equation}
 w\left(r\right)=\sum_{l\in \mathbb{Z}}\alpha_{j_0,l} \underbrace{\phi_{j_0,l}\left(r\right)}_{\text{scaling functions}} + \sum_{j\geq j_0}^{\infty} \sum_{l \in \mathbb{Z}}\beta_{j,l} \underbrace{\psi_{j,l}\left(r\right)}_{\text{wavelets}}, 
\label{eq:KernelExpansion}
\end{equation}
where $\alpha_{j_0,l}$ are the approximation coefficients at the lowest scale $j_0$, and $\beta_{j,l}$ are the detail coefficients at different scales $j$, with
\begin{align}
\phi_{j,l}\left(r\right) =& 2^{\frac{j}{2}}\phi\left(2^j r-l\right) \\	
\psi_{j,l}\left(r\right) =& 2^{\frac{j}{2}}\psi\left(2^j r-l\right).
\end{align}
The parameters $j$ and $l$ control the scale (dilation) and translation (spatial shift), respectively. \dean{NEED SOMETHING ABOUT HOW j AND l ARE SET HERE - AS IN THAT THEY ARE KNOWN - AND THAT WE ONLY NEED TO ESTIMATE THE COEFFICIENTS.} \parham{The number of spatial shifts is dependent on the scale, $j$, and can be chosen such that the basis functions cover the spatial domain of interest. The number of scales depends on the spectral characteristics of the field, which will be described later.} Scaling functions retain the lowest frequency components, due to their low-pass properties, while the wavelet functions extract successively higher frequency components due to their band-pass characteristics.

@@

The $m$-th  order cardinal B-spline function is defined by the recurrence relation \citep{Chui1992} 
\begin{equation}
N_{m}\left(r\right) = \left(N_{m-1}\ast N_{1}\right)\left(r\right) = \int_0^{1} N_{m-1}\left( r-r'\right)\,\mathrm{d}r',
\label{SplineConvolutionIntegral}
\end{equation}
where $m>1$, $\ast$ denotes convolution, and $N_1\left(r\right)$ is the characteristic function of the unit interval $\left[ 0,1\right)$, also known as indicator function, i.e.
\begin{equation}
N_{1}\left(r\right)=
\begin{cases}
1 & \text{if $ 0\le r<1$}, \\
0 & \mathrm{elsewhere}.
\end{cases}
\end{equation}
Following this, a B-spline of any order, $N_m(r)$, can be computed using the recursive expression defined by \citep{DeBoor2001}
\begin{equation}\label{eq:MRA-DoBoorFormula}
 N_{m}\left(r\right)=\frac{r}{m-1}N_{m-1}\left(r\right)+\frac{m-r}{m-1}N_{m-1}\left(r-1\right) \quad m>1.
 \end{equation}
The most commonly used form of splines is the cubic spline $\left(m=4\right)$, comprised of third degree polynomials added together at the joining points, with an explicit expression given by
\begin{align}
3!N_{4}\left(r\right)=
\begin{cases}
r^3 & \text{if $ 0\le r\le1$}, \\
4-12r+12r^2-3r^3 & \text{if $1\le r\le2$}, \\
-44+60r-24r^2+3r^3 & \text{if $2\le r\le3$}, \\
64-48r+12r^2-r^3 & \text{if $3\le r\le4$}, \\
0 & \mathrm{elsewhere}.
\end{cases}
\end{align}
One important feature of the B-spline functions is the \emph{partition of unity} property, meaning that the unity can be expressed as a linear sum of B-splines, i.e
\begin{equation}
	\sum_{l}N_m(r+l)\equiv1
	\end{equation}
	\parham{This is important as the underlying field can be reconstructed when its saturated without introducing numerical errors which is not the case for example in the case of Gaussian basis functions.}

@@

The multi-resolution approximation using B-spline functions is completed by defining a two-scale relation pair, which relates the scaling functions and the wavelets at a given scale with the scaling function at the next higher scale. For cardinal B-spline scaling and wavelet functions of order $m$ the two-scale relation pair takes the form of
\begin{align}
 N_{m}\left(r\right)&=\sum_{n=0}^{m} \left[\mathbf p\right]_n N_{m}\left(2r-n\right) \label{eq:MRA-TwoScalepair1} \\
  \varphi_{m}\left(r\right) &= \sum_{n=0}^{3m-2} \frac{\left(-1\right)^n}{2^{m-1}} \left[\mathbf q\right]_n N_{m}\left(2r-n\right)\label{eq:MRA-TwoScalepair2},
 \end{align}
where 
 \begin{align}
\left[\mathbf p\right]_n&=2^{-m+1} \binom{m}{n} \quad \text{ $0\le n\le m$} \label{eq:MRA-TwoScalepair1coefs}\\
\left[\mathbf q\right]_n&= \sum_{l=0}^{m} \binom{m}{l} N_{2m}\left(n-l+1\right), \,  \text{ $0\le n\le 3m-2$}\label{eq:MRA-TwoScalepair2coefs}
 \end{align}
and $\left[\cdot\right]_n$ denotes the $n$-th element of the vector coefficients \citep{Chui1992}. \dean{Need to explicitly say why this is important and what this gives us here.}

@@

\begin{align}\label{eq:Xi0derivation0} 
 \mathbf M_{t+1}&=\mathbf E_{\theta'}\left[(\mathbf x_t-\mathbf{\hat x}_t)(\mathbf x_{t+1}-\mathbf{\hat x}_{t+1})^\top\right] \nonumber \\
&=\mathbf E_{\theta'}\left[\mathbf x_t\mathbf x_{t+1}^\top\right]-\mathbf E_{\theta'}\left[\mathbf x_t\hat{\mathbf x}_{t+1}^\top\right]-\mathbf E_{\theta'}\left[\hat{\mathbf x}_t\mathbf x_{t+1}^\top\right]+\mathbf E_{\theta'}\left[\hat{\mathbf x}_t\hat{\mathbf x}_{t+1}^\top\right]  \nonumber \\
&= \mathbf E_{\theta'}\left[\mathbf x_t\mathbf x_{t+1}^\top\right]-\hat{\mathbf x}_t\hat{\mathbf x}_{t+1}^\top-\hat{\mathbf x}_t\hat{\mathbf x}_{t+1}^\top+\hat{\mathbf x}_t\hat{\mathbf x}_{t+1}^\top\nonumber \\
&= \mathbf E_{\theta'}\left[\mathbf x_t\mathbf x_{t+1}^\top\right]-\hat{\mathbf x}_t\hat{\mathbf x}_{t+1}^\top.
\end{align}
Rearranging  equation \eqref{eq:Xi0derivation0} we get
\begin{equation}\label{eq:Xi0derivation1}
 \mathbf E_{\theta'}\left[\mathbf x_t\mathbf x_{t+1}^\top\right]=\mathbf M_{t+1}+\mathbf {\hat x}_t\mathbf{\hat x}_{t+1}^\top
\end{equation}
substituting \ref{eq:Xi0derivation1} into \ref{eq:appXi0}, noting the expectation commutes with the sum we get
\begin{equation}
 \boldsymbol\Xi_1=\sum_{t=0}^{T-1}\left(\mathbf M_{t+1}+\mathbf{\hat x}_t\mathbf{\hat x}_{t+1}^\top\right)
\end{equation}   
An expression for $\boldsymbol\Xi_1$ in terms of the smoothed states can be also found in a similar way by expanding the covariance matrix,
\begin{equation}\label{eq:appXi1}
 \boldsymbol\Xi_1=\mathbf E_{\theta'}\left[\sum_{t=0}^{T-1}\mathbf x_{t}\mathbf x_{t}^\top\right],
\end{equation}
therefore we can write
\begin{equation}\label{eq:Xi1derivation1}
 \mathbf E_{\theta'}\left[\mathbf x_{t}\mathbf x_{t}^\top\right]=\mathbf P_{t}+\mathbf {\hat x}_{t}\mathbf{\hat x}_{t}^\top,
\end{equation}
substituting \ref{eq:Xi1derivation1} into \ref{eq:appXi1} we get
\begin{equation}
 \boldsymbol\Xi_1=\sum_{t=0}^{T-1}\left(\mathbf P_{t}+\mathbf{\hat x}_{t}\mathbf{\hat x}_{t}^\top\right).
\end{equation}
\newpage
\bibliographystyle{./elsarticle/model2-names} 
\bibliography{MRAIDE}

>>>>

 % TO DELETE

%    \documentclass[journal,a4paper]{IEEEtran}
%  \documentclass[journal]{IEEEtran}

%     \documentclass[10pt,twocolumn,twoside]{IEEEtran}
     \documentclass[11pt,draftcls,onecolumn,peerreview]{IEEEtran} 
%         \topmargin       -6.0mm
%          \oddsidemargin      0mm
%          \evensidemargin     0mm
%          \textheight     223.5mm
%         \textwidth      170.0mm
\usepackage{setspace} 
\usepackage{color}
\newcommand{\mike}[1]{\textcolor{red}{#1}}
\newcommand{\dean}[1]{\textcolor{green}{#1}}
\newcommand{\parham}[1]{\textcolor{blue}{#1}} 
\newcommand{\ken}[1]{\textsf{\emph{\textbf{\textcolor{magenta}{#1}}}}} 
\newcommand{\cut}[1]{\textcolor{cyan}{#1}}

\usepackage{cite}
%Reduce the spacing between figures and captions
% \usepackage[belowskip=-15pt,aboveskip=10pt]{caption}
%  \setlength{\intextsep}{2pt plus 2pt minus 2pt}

\ifCLASSINFOpdf
   \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi

\usepackage[cmex10]{amsmath}
\interdisplaylinepenalty=2500
\usepackage{amssymb}
\ifCLASSOPTIONcompsoc
  \usepackage[tight,normalsize,sf,SF]{subfigure}
\else
  \usepackage[tight,footnotesize]{subfigure}
\fi

\usepackage{natbib}

\usepackage{stfloats}
\usepackage{float}
\floatstyle{ruled}
%This bit is how to use hyperlink without stupid coloured rectangles around it
\usepackage{float}
\usepackage{hyperref}
\hypersetup{colorlinks,linkcolor=black,filecolor=black,urlcolor=black,citecolor=black}
%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
\newfloat{algorithm}{htbp}{loa}%[chapter]
\floatname{algorithm}{Algorithm}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model}

\author{P. Aram, D.R. Freestone, M. Dewar, K.Scerri, D.B. Grayden and V. Kadirkamanathan % <-this % stops a space
\thanks{P. Aram* and V. Kadirkamanathan are with the Department of Automatic Control and Systems Engineering, University of Sheffield, Sheffield, S1 3JD, U.K. (e-mail:parham.aram@univ-amu.fr; visakan@sheffield.ac.uk).}% <-this % stops a space
\thanks{D.\ R.\ Freestone and D.\ B.\ Grayden are with the Department
of Electrical and Electronic Engineering, The University of Melbourne, and The Bionic Ear Institute, VIC, Australia. (e-mail:dfreestone@bionicear.org; grayden@unimelb.edu.au). The Bionic Ear Institute acknowledges the support it receives from the Victorian State Government through the Operational Infrastructure Support Program.}
\thanks{M. Dewar is with the Department of Applied Physics and Applied Mathematics, Columbia University, New York, NY, USA. (e-mail:mike.dewar@columbia.edu).}
\thanks{K. Scerri is with the Department of Systems and Control Engineering, University
of Malta, Msida, MSD, Malta. (e-mail:kenneth.scerri@um.edu.mt).}}
% <-this % stops a space
%\thanks{Manuscript received April 19, 2005; revised January 11, 2007.}}

% The paper headers
% \markboth{Journal of }%
% {Aram \MakeLowercase{\textit{et al.}}: Wavelet Multiresolution Spatio-Temporal Modelling Using the Integro-Difference Equation}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.
 \ifCLASSOPTIONpeerreview
\else
\fi

% If you want to put a publisher's ID mark on the page you can do it like
% this:
%\IEEEpubid{0000--0000/00\$00.00~\copyright~2007 IEEE}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

% make the title area
\maketitle

\begin{abstract}
%\boldmath
We develop a multi-resolution approximation (MRA) framework for the integro-difference equation (IDE) neural field model based on semi-orthogonal cardinal B-spline wavelets. State and parameter estimation is performed using the Expectation Maximization (EM) algorithm. A synthetic example is provided
to demonstrate the framework.
\end{abstract}
% Note that keywords are not normally used for peerreview papers.
 \begin{IEEEkeywords}
  Neural field model, multi-resolution approximation (MRA), Expectation Maximization (EM) algorithm, wavelets.
  \end{IEEEkeywords}
% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
%   \begin{center} \bfseries EDICS Category: SSP-IDEN \end{center}
%   \fi
%
%  For peerreview papers, this IEEEtran command inserts a page break and
%creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
\IEEEPARstart{T}{he} human cerebral cortex has a multi-resolution architecture, where spatial scales for information processing range from ion channels, to single neurons, to networks of millions of neurons. This multi-resolution cortical architecture poses major modeling challenges to efficiently describe the brain's dynamics. This paper introduces a multi-resolution data-driven framework for neural field modeling, the Multi-Resolution Approximate Integro-Difference Equation (MRAIDE), to address this challenge.

@@

\section{IDE Neural Field Model}
\singlespacing
\begin{table*}[!t]
\begin{tabular}{|l|l|l|}
	\hline
	\textbf{Symbol} & \textbf{Quantity} & \textbf{Units}\\
	\hline
	\multicolumn{3}{|c|}{\emph{Domain and indices}}\\
	\hline
	$\Omega$ & spatial domain & n.a.\\
	$\mathbf{r}$ & spatial location & [mm, mm]\\
	$t$ & time & s\\
	\hline
	\multicolumn{3}{|c|}{\emph{Model}}\\
	\hline
    $y_t$ & observation & mV\\
    $v(\mathbf{r},t)$ & mean membrane potential & mV \\
	$f(v\left(\mathbf{r},t\right))$ & firing rate function & spikes.s$^{-1}$\\
	$e(\mathbf{r},t)$ & field disturbance, with covariance function $\eta(\mathbf r-\mathbf r')$ & mV\\
	$\boldsymbol\varepsilon_t$ & observation noise, with covariance matrix $\Sigma_\varepsilon$ & mV\\
	$m(\mathbf{r}_n)$ & observation kernel where $n$ is sensor index $n=1,..,n_y$ & n.a. \\
	$w(\mathbf{r})$ & spatial connectivity kernel & n.a.\\
	$h(t)$ & post-synaptic response kernel & mV\\
	$\tau$, $\xi$ & \parham{synaptic time constant} \& time constant parameter & s, n.a.\\
	$\delta(t)$ & Dirac-delta function & n.a.\\
	\hline
	\multicolumn{3}{|c|}{\emph{Reduced Model}} \\
	\hline
   	$\mathbf{x}_t$ & state vector at time $t$ & mV\\
   	$\mathbf{e}_t$ & state disturbance vector, with covariance $\Sigma_e$ & mV\\
   	$\mathbf{C}$ & observation matrix & n.a. \\
	\hline
	\multicolumn{3}{|c|}{\emph{Frequency Analysis}} \\
	\hline
	$\boldsymbol{\nu}$, $\boldsymbol{\nu}_c$ & spatial frequency and spatial cutoff frequency & cycles/mm \\
	$\sigma_{\nu}^2$ & variance of Fourier transformed Gaussian basis function & (cycles/mm)$^2$\\
	\hline
	\multicolumn{3}{|c|}{\emph{Estimation}} \\
	\hline
	$\hat{\mathbf{x}}_t^{f-}$, $\hat{\mathbf{x}}_t^f$ & forward prior and posterior state estimates & mV\\
	$\hat{\mathbf{x}}_t^{b}$ & backward posterior state estimates & mV\\
	$P^{f-}_t$, $P^f_t$  & forward prior and posterior covariance matrices & mV$^2$\\
	$P^b_t$ & backward posterior covariance matrices & mV$^2$\\
	$M_t$& cross-covariance matrix & mV$^2$\\
	$\mathcal K_{t} $ & filter gain & n.a.\\ 
	$\mathcal S_{t} $ & smoother gain & n.a.\\
	\hline
\end{tabular}
\caption{\textbf{Notation.} The first column of the table gives the symbols for all the notation used in the paper. The second column provides a brief description of the quantity, and the third column provides the SI units where relevant.}
\label{tab:Notation}
\end{table*}
\doublespacing

@@

The multi-resolution approximation \citep{Mallat1989a} of the neural IDE (MRAIDE) is obtained by decomposing both the field, $v_t(\cdot)$, and the connectivity kernel, $w(\cdot)$, (assuming square-integrable functions) using translations and dilations of a scaling function $\phi(\mathbf{r})$ and a mother wavelet $\psi(\mathbf{r})$. Considering a  one-dimensional field, the static connectivity kernel is decomposed as,
\begin{equation}
 w\left(r\right)=\sum_{l\in \mathbb{Z}}\alpha_{j_0,l} \underbrace{\phi_{j_0,l}\left(r\right)}_{\text{scaling functions}} + \sum_{j\geq j_0}^{\infty} \sum_{l \in \mathbb{Z}}\beta_{j,l} \underbrace{\psi_{j,l}\left(r\right)}_{\text{wavelets}}, 
\label{eq:KernelExpansion}
\end{equation}
where $\alpha_{j_0,l}$ are the approximation coefficients at the lowest scale $j_0$, and $\beta_{j,l}$ are the detail coefficients at different scales $j$, with
\begin{align}
\phi_{j,l}\left(r\right) =& 2^{\frac{j}{2}}\phi\left(2^j r-l\right) \\	
\psi_{j,l}\left(r\right) =& 2^{\frac{j}{2}}\psi\left(2^j r-l\right).
\end{align}
The parameters $j$ and $l$ control the scale (dilation) and translation (spatial shift), respectively. \dean{NEED SOMETHING ABOUT HOW j AND l ARE SET HERE - AS IN THAT THEY ARE KNOWN - AND THAT WE ONLY NEED TO ESTIMATE THE COEFFICIENTS.} \parham{The number of spatial shifts, $l$, is dependent on the scale, $j$, and can be chosen such that the basis functions cover the spatial domain of interest. The number of scales depends on the spectral characteristics of the field, which will be described later.} Scaling functions retain the lowest frequency components, due to their low-pass properties, while the wavelet functions extract successively higher frequency components due to their band-pass characteristics.

@@

The $m$-th  order cardinal B-spline function is defined by the recurrence relation \citep{Chui1992} 
\begin{equation}
N_{m}\left(r\right) = \left(N_{m-1}\ast N_{1}\right)\left(r\right) = \int_0^{1} N_{m-1}\left( r-r'\right)\,\mathrm{d}r',
\label{SplineConvolutionIntegral}
\end{equation}
where $m>1$, $\ast$ denotes convolution, and $N_1\left(r\right)$ is the characteristic function of the unit interval $\left[ 0,1\right)$, also known as indicator function, i.e.
\begin{equation}
N_{1}\left(r\right)=
\begin{cases}
1 & \text{if $ 0\le r<1$}, \\
0 & \mathrm{elsewhere}.
\end{cases}
\end{equation}
Following this, a B-spline of any order, $N_m(r)$, can be computed using the recursive expression defined by \citep{DeBoor2001}
\begin{equation}\label{eq:MRA-DoBoorFormula}
 N_{m}\left(r\right)=\frac{r}{m-1}N_{m-1}\left(r\right)+\frac{m-r}{m-1}N_{m-1}\left(r-1\right) \quad m>1.
 \end{equation}
The most commonly used form of splines is the cubic spline $\left(m=4\right)$, comprised of third degree polynomials added together at the joining points, with an explicit expression given by
\begin{align}
3!N_{4}\left(r\right)=
\begin{cases}
r^3 & \text{if $ 0\le r\le1$}, \\
4-12r+12r^2-3r^3 & \text{if $1\le r\le2$}, \\
-44+60r-24r^2+3r^3 & \text{if $2\le r\le3$}, \\
64-48r+12r^2-r^3 & \text{if $3\le r\le4$}, \\
0 & \mathrm{elsewhere}.
\end{cases}
\end{align}
One important feature of the B-spline functions is the \emph{partition of unity} property, meaning that the unity can be expressed as a linear sum of B-splines, i.e
\begin{equation}
	\sum_{l}N_m(r+l)\equiv1
	\end{equation}
	This is important as the underlying field can be reconstructed when its saturated without introducing numerical errors which is not the case for example in the case of Gaussian basis functions.

@@

The multi-resolution approximation using B-spline functions is completed by defining a two-scale relation pair, which relates the scaling functions and the wavelets at a given scale with the scaling function at the next higher scale. This is important for constructing the wavelets. For cardinal B-spline scaling and wavelet functions of order $m$ the two-scale relation pair takes the form of
\begin{align}
 N_{m}\left(r\right)&=\sum_{n=0}^{m} \left[\mathbf p\right]_n N_{m}\left(2r-n\right) \label{eq:MRA-TwoScalepair1} \\
  \varphi_{m}\left(r\right) &= \sum_{n=0}^{3m-2} \frac{\left(-1\right)^n}{2^{m-1}} \left[\mathbf q\right]_n N_{m}\left(2r-n\right)\label{eq:MRA-TwoScalepair2},
 \end{align}
where 
 \begin{align}
\left[\mathbf p\right]_n&=2^{-m+1} \binom{m}{n} \quad \text{ $0\le n\le m$} \label{eq:MRA-TwoScalepair1coefs}\\
\left[\mathbf q\right]_n&= \sum_{l=0}^{m} \binom{m}{l} N_{2m}\left(n-l+1\right), \,  \text{ $0\le n\le 3m-2$}\label{eq:MRA-TwoScalepair2coefs}
 \end{align}
and $\left[\cdot\right]_n$ denotes the $n$-th element of the vector coefficients \citep{Chui1992}. \dean{Need to explicitly say why this is important and what this gives us here.}

@@

\begin{align}\label{eq:Xi0derivation0} 
 \mathbf M_{t+1}&=\mathbf E_{\theta'}\left[(\mathbf x_t-\mathbf{\hat x}_t)(\mathbf x_{t+1}-\mathbf{\hat x}_{t+1})^\top\right] \nonumber \\
&=\mathbf E_{\theta'}\left[\mathbf x_t\mathbf x_{t+1}^\top\right]-\mathbf E_{\theta'}\left[\mathbf x_t\hat{\mathbf x}_{t+1}^\top\right]-\mathbf E_{\theta'}\left[\hat{\mathbf x}_t\mathbf x_{t+1}^\top\right]+\mathbf E_{\theta'}\left[\hat{\mathbf x}_t\hat{\mathbf x}_{t+1}^\top\right]  \nonumber \\
&= \mathbf E_{\theta'}\left[\mathbf x_t\mathbf x_{t+1}^\top\right]-\hat{\mathbf x}_t\hat{\mathbf x}_{t+1}^\top-\hat{\mathbf x}_t\hat{\mathbf x}_{t+1}^\top+\hat{\mathbf x}_t\hat{\mathbf x}_{t+1}^\top\nonumber \\
&= \mathbf E_{\theta'}\left[\mathbf x_t\mathbf x_{t+1}^\top\right]-\hat{\mathbf x}_t\hat{\mathbf x}_{t+1}^\top.
\end{align}
Rearranging  equation \eqref{eq:Xi0derivation0} we get
\begin{equation}\label{eq:Xi0derivation1}
 \mathbf E_{\theta'}\left[\mathbf x_t\mathbf x_{t+1}^\top\right]=\mathbf M_{t+1}+\mathbf {\hat x}_t\mathbf{\hat x}_{t+1}^\top
\end{equation}
substituting \ref{eq:Xi0derivation1} into \ref{eq:appXi0}, noting the expectation commutes with the sum we get
\begin{equation}
 \boldsymbol\Xi_1=\sum_{t=0}^{T-1}\left(\mathbf M_{t+1}+\mathbf{\hat x}_t\mathbf{\hat x}_{t+1}^\top\right)
\end{equation}   
An expression for $\boldsymbol\Xi_1$ in terms of the smoothed states can be also found in a similar way by expanding the covariance matrix,
\begin{equation}\label{eq:appXi1}
 \boldsymbol\Xi_1=\mathbf E_{\theta'}\left[\sum_{t=0}^{T-1}\mathbf x_{t}\mathbf x_{t}^\top\right],
\end{equation}
therefore we can write
\begin{equation}\label{eq:Xi1derivation1}
 \mathbf E_{\theta'}\left[\mathbf x_{t}\mathbf x_{t}^\top\right]=\mathbf P_{t}+\mathbf {\hat x}_{t}\mathbf{\hat x}_{t}^\top,
\end{equation}
substituting \ref{eq:Xi1derivation1} into \ref{eq:appXi1} we get
\begin{equation}
 \boldsymbol\Xi_1=\sum_{t=0}^{T-1}\left(\mathbf P_{t}+\mathbf{\hat x}_{t}\mathbf{\hat x}_{t}^\top\right).
\end{equation}
\ifCLASSOPTIONcaptionsoff
  \newpage
\fi

%   \newpage
% \bibliographystyle{unsrt} 
\bibliographystyle{apalike}
\bibliography{MRAIDE}

% \bibliographystyle{IEEEtran}
% argument is your BibTeX string definitions and bibliography database(s)
 % \bibliography{IEEEabrv,MRAIDE}
